\documentclass{sigplanconf}

\usepackage{amsmath}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Generating LilyPond Code in Python via Abjad to Create Music Notation Documents}
%\subtitle{Subtitle Text, if any}

\authorinfo{Josiah Wolf Oberholtzer}
           {Harvard University}
           {josiah.oberholtzer@gmail.com}
\authorinfo{Jeffrey Trevi\~{n}o}
           {Colorado College}
           {jeffrey.trevino@gmail.com}
\authorinfo{Trevor Ba\v{c}a}
           {Harvard University}
           {trevor.baca@gmail.com}

\maketitle

\begin{abstract}
LilyPond is a modular, extensible, programmable music notation document
compiler that converts text input into automatically formatted graphic output.
The Abjad API for Formalized Score Control -- which extends the Python
programming language with an open-source, object-oriented model of
common-practice music notation that enables users to build music notation
documents through the aggregation of elemental music notation objects --
generates well-formed LilyPond input syntax. Abjad models hierarchically
organized music information as a reference-rich, normalized tree. In order to
model musical-typographical constructs such as slurs, beams and ties, Abjad
introduces cyclicity into its core tree data structure via bidirectionally
referenced, subtree-spanning nodes. Abjad traverses this data structure to
collect formatting information from each encountered node, and a hybrid set of
rules derived from graph theory, western music notation, and the constraints of
LilyPond's domain-specific language enables Abjad to organize this collected
information into a LilyPond input string.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

\terms
term1, term2

\keywords
keyword1, keyword2

\section{Introduction}

The text of the paper begins here.

\section{LilyPond syntax and data model}
Lilypond compiles music documents via two distinct module. First, a module parses the input
file, and then a module formats the music for tyesetting. The compiler translates the parsed text input into an event stream representation before formatting to streamline music typesetting process.\cite{sandberg2006separating}

Lilypond requires grammar rules that are shaped in a certain rules (certain commands can be in certain locations with greater or lesser degrees of flexibility). Pitch name, followed by the octave tick (or lack thereof), and the duration.

Lilypond's context hierarchy. Lilypond models music notation as a hierarchy of contexts, each of which is a collection of engravers, modules that handle specific typographical tasks. Lilypond's modular engraver design entails that well-formed element grammar commands apply to specific contexts and not to other.

Lilypond is an interpreted language. It's stateful.

\section{Abjad data model}
While Lilypond implies containers and spanners through the interaction of the text input and event stream representations, Abjad models containers and spanners explicitly.

\section{Traversal, collection and organization}

\subsection{Abjad's format manager}

\subsection{Format bundles and lexical organization of format contributions}

\section{Discrepancies between Abjad and LilyPond}

\subsection{Parsing LilyPond syntax in Abjad}

\section{Conclusion}

\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright

\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
P. Q. Smith, and X. Y. Jones. ...reference text...

\end{thebibliography}


\end{document}
